---
title: "p8105_hw1"
author: "Hengxuan Ma"
date: "2022-09-23"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
```

# Problem 1

## Load the dataset

Load the penguins dataset.

```{r}
data("penguins", package = "palmerpenguins")
```

## Write a short description of the dataset

The penguins dataset has `r ncol(penguins)` important variables: `r colnames(penguins)`. 
There are `r nlevels(pull(penguins, species))` levels under the variable `r colnames(penguins)[1]`.
There are `r nlevels(pull(penguins, island))` levels under the variable `r colnames(penguins)[2]`.
The variable `r colnames(penguins)[8]` has values of `r unique(pull(penguins, year))`.
The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r mean(as.numeric(pull(penguins, flipper_length_mm)), na.rm = TRUE)`.

## Make a scatterplot

Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x).

```{r warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()
```

Export the scatterplot to the project directory.

```{r warning = FALSE}
ggsave("scatter_plot.pdf")
```


