---
title: "p8105_hw1"
author: "Hengxuan Ma"
date: "2022-09-23"
output: github_document
---

```{r include = FALSE}
library(tidyverse)
```

# Problem 1

## Load the dataset

Load the penguins dataset.

```{r}
data("penguins", package = "palmerpenguins")
```

## Write a short description of the dataset

The penguins dataset has `r ncol(penguins)` important variables: `r colnames(penguins)`. 
There are `r nlevels(pull(penguins, species))` levels under the variable `r colnames(penguins)[1]`.
There are `r nlevels(pull(penguins, island))` levels under the variable `r colnames(penguins)[2]`.
The variable `r colnames(penguins)[8]` has values of `r unique(pull(penguins, year))`.
The dataset has `r nrow(penguins)` rows and `r ncol(penguins)` columns. 
The mean flipper length is `r mean(as.numeric(pull(penguins, flipper_length_mm)), na.rm = TRUE)`.

## Make a scatterplot

Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x).

```{r warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()
```

Export the scatterplot to the project directory.

```{r warning = FALSE}
ggsave("scatter_plot.pdf")
```

# Problem 2

## Create a data frame

Create a data frame comprised of different variables.
```{r}
x = rnorm(10, sd = 1)
p2_df = tibble(
  vec_numeric = x,
  vec_logic = x > 0,
  vec_char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  vec_factor = factor(c("level1", "level2", "level1", "level3", "level2", "level1", "level2", "level3", "level3", "level1"))
)
```

## Try to take the mean of each variable in the dataframe

Take the mean of numeric variables.

```{r}
df_num = pull(p2_df, vec_numeric)
mean(df_num)
```

Take the mean of logical variables.

```{r}
df_logic = pull(p2_df, vec_logic)
mean(df_logic)
```

Take the mean of character variables.

```{r warning = FALSE}
df_char = pull(p2_df, vec_char)
mean(df_char)
```

Take the mean of factor variables.

```{r warning = FALSE}
df_factor = pull(p2_df, vec_factor)
mean(df_factor)
```

Therefore, the mean of numeric and logical variables can be successfully calculated. But the character and factor variables don't work.

## Convert variables from one type to another

Applies the as.numeric function to the logical variables.

```{r}
logic_convert = as.numeric(df_logic)
logic_convert
```

Applies the as.numeric function to the character variables.

```{r warning = FALSE}
char_convert = as.numeric(df_char)
char_convert
```

Applies the as.numeric function to the factor variables.

```{r}
factor_convert = as.numeric(df_factor)
factor_convert
```

According to the results from as.numeric function, we can see that logical and factor variables can be converted into numbers while character variables can not. Since values of logical variables are binary (TRUE=1 and FALSE=0), the mean of logical variables can be taken. Although values of factor variables can be converted to the number of levels, the mean of factor variables cannot be directly taken. However, the value of character variable consists only of alphabets and has no relationship with numbers, it cannot be converted into the numeric form and we cannot take the mean.




